### 2021-03-15 notes/notes.tex
Added Cue and Clamp to both task and models.

In this experiment, every context $C_i$ is characterized by the strength and
direction of its force: $C_i = \{\theta_i, \gamma_i\}$, where $i = 1, 2, ...,
N_c$, $N_c$ is the number of contexts, $\theta$ is the angle of the force
measured Cartesianly, and $w_i$ is the intensity of the force.

@@

Once the posterior over contexts has been calculated for the current trial, it
is used to select the current context. A simple rule is to sample the context
directly from the posterior distribution. However, a useful extension of this
is to use a softmax function to create a distribution:

\begin{equation}
S(C^{(t)}) = \frac{e^{\kappa
q\left(C^{(t)}\right)}}{\displaystyle \sum_{i=1}^{N_c}e^{\kappa
q\left(C_i\right)}}
\end{equation}
where $q(C_i)$ is given by \eref{eqn:estimated-context}. The
function $S(\cdot)$ is then used to sample the (inferred) current context. This
representation has the advantage of adding a parameter with which the sampling
behavior can be characterized, ranging from choosing uniformly when $\kappa =
0$ to choosing the context with the highest probability deterministically as
$\kappa \rightarrow \infty$.

@@

For simplicity, we assume that participants know how many distinct contexts
will be shown to them, and therefore all the $C_x$ have been created by the
agent before the session begins. By default, all contexts assume a force with
angle and magnitude equal zero, and their true values are learned during the
experiment (see below).

For mathematical tractability (i.e. to avoid circular distributions), we will
limit contexts to one where the force points left (angle $\pi$) and one where
the force points right (angle 0), in addition to the baseline context. In this
scheme, participants need only learn the magnitude of the force for each
context. We will assume that for each context, the agent's belief over the
magnitude of the force is given by a normal distribution:
\begin{equation}
p(\gamma) = N(\mu_f, \sigma_f) \label{eqn:data-dist}
\end{equation}
where $\mu_f$ and $\sigma_f$ are parameters to be estimated at each
trial. Before the experiment begins, the agent will have a prior distribution
over these parameters. A standard Bayesian approach is to choose NormalGamma
priors:
\begin{equation}
p(\mu_f, \sigma_f) = NG(\mu_0, \nu_0, \alpha_0, \beta_0)
\end{equation}
where $\mu_0, \nu_0, \alpha_0$ and $\beta_0$ are free hyperparameters of the
model. Because participants are told that there are three contexts (baseline,
left and right), we assume the following hyperparameter values:
\begin{align} \mu_f^{(0)} &= (0, -1, 1) \\
  \nu_f^{(0)} &= 0.1 \\
  \alpha_f^{(0)} &= 1 \\
  \beta_f^{(0)} &= 0.5
\end{align}
Note that $\mu_f^{(0)}$ is different for each context and is thus
provided as a vector $(0, -1, 1)$ representing the baseline, left and right
contexts, respectively. For all other hyperparameters, all contexts have the
same value. These values ensure that the prior over the hyperparameters
$(\mu_f, \sigma_f)$ has the mode at $(\mu_f^{(0)}, 1)$. We will show that the
choice of these hyperparameters affects learning rate, but not the final
learned magnitude \todo{Check this}.

@@

The final step is action selection. Action selection is done at the level of
desired outcome (as opposed to motor commands) following the equations:
\begin{align}
  x_{t+1} &= x_t + a_t + f_t + \sqrt{\Delta
            t}\epsilon\\ \label{eqn:dynamics}
  f_t &= \gamma_t (cos\theta_t, sin\theta_t) \\
  a_t &= \argmax_{a_t}p(x_{t+1} = (0, 0) | x_t, a_t, f_t)
\end{align}
where $\gamma_t$ and $\theta_t$ are the participant's estimate of
the parameters, given by \eref{eqn:estimated-context}. $\epsilon$ is a Gaussian
motor error component.

>>>>

In this experiment, every context $C_i$ is characterized by the strength and
8'rection of its force: $C_i = \{\theta_i, \gamma_i\}$, where $i = 1, 2, ...,
N_c$, $N_c$ is the number of contexts, $\theta$ is the angle of the force
measured Cartesianly, and $w_i$ is the intensity of the force.

@@

Once the posterior over contexts has been calculated for the current trial, it
is used to select the current context. A simple rule is to sample the context
directly from the posterior distribution. However, a useful extension of this
is to use a softmax function to create a distribution:
 !''¿
\begin{equation}
S(C^{(t)}) = \frac{e^{\kappa
q\left(C^{(t)}\right)}}{\displaystyle \sum_{i=1}^{N_c}e^{\kappa
q\left(C_i\right)}}
\end{equation}
where $q(C_i)$ is given by \eref{eqn:estimated-context}. The
function $S(\cdot)$ is then used to sample the (inferred) current context. This
representation has the advantage of adding a parameter with which the sampling
behavior can be characterized, ranging from choosing uniformly when $\kappa =
0$ to choosing the context with the highest probability deterministically as
$\kappa \rightarrow \infty$.

@@

  For simplicity, we assume that participants know how many distinct contexts
will be shown to them, and therefore all the $C_x$ have been created by the
agent before the session begins. By default, all contexts assume a force with
angle and magnitude equal zero, and their true values are learned during the
experiment (see below).
1234568¿|
For mathematical tractability (i.e. to avoid circular distributions), we will
limit contexts to one where the force points left (angle $\pi$) and one where
the force points right (angle 0), in addition to the baseline context. In this
scheme, participants need only learn the magnitude of the force for each
context.

The generative model for the force can take on different shapes for the prior
and posterior distributions. In this work, we explore three different forms of
this generative model which differ on the assumptions the agent makes regarding
how the force fields are being generated by the environment. Each version of
the model is discussed separately below and their pros and cons discussed;
additionally, simulations are presented in the following sections.

To infer the parameters of the posterior distributions, we make use of Bayes'
theorem. In the following equations, we drop the dependence on the context and
observations for notational simplicity:
\begin{equation}
q(\theta | x_{1:t}) \propto p(x_{1:t} | \theta)p(\beta)
\end{equation}
where $\theta$ are the parameters of the force field to be estimated and
$x_{1:t}$ are the data observed until time $t$. Across all variants of the
model presented below, the likelihood will be a Gaussian:
\begin{align}
  p(x_t | \theta) &= N(\mu_x, \sigma_x)  \\
  \mu_x &= \theta_{t-1} + x \\
  \sigma_x &= \xi_0 1
\end{align}
This likelihood is the probability of observing $x$ given the previous estimate
of the parameters of the force field $\theta_{t-1}$, and given the dynamics of
the system (see below).

The final part of the model to establish is action selection. Action selection
is done at the level of desired outcome (as opposed to motor commands)
following the equations:
\begin{align}
  x_{t+1} &= x_t + a_t + f_t + \sqrt{\Delta
            t}\epsilon\\ \label{eqn:dynamics}
  f_t &= \gamma_t (cos\theta_t, sin\theta_t) \\
  a_t &= \argmax_{a_t}p(x_{t+1} = (0, 0) | x_t, a_t, f_t)
\end{align}
where $\gamma_t$ and $\theta_t$ are the participant's estimate of
the parameters, given by \eref{eqn:estimated-context}. $\epsilon$ is a Gaussian
motor error component.

\subsection{Fixed force value}
In this version, the force field is assumed to have a fixed value which the
agent tries to infer. The agent's uncertainty on this inference is reflected on
the parameters of the posterior over the force directly. For simplicity, we
assume that the priors (and posteriors) are normal with mean and standard
deviation to be updated at every time step. Therefore, the posterior over the
magnitude of the force for each context is given by:
\begin{equation}
q_t(\gamma) = N(\mu_{\gamma}, \sigma_\gamma)
\end{equation}
where $q_t(\gamma)$ is the posterior of the magnitude $\gamma$ of the force
field distribution after the observations at time $t$, and $\mu_\gamma$ and
$\sigma_\gamma$ are the inferred parameters of the distribution.

Because the likelihood is a Gaussian with support $\mu_\gamma$, the update
equations are given by:
\begin{align}
  \mu_\gamma^{(t)} &= \frac{\mu_\gamma^{(t-1)}\left(\sigma_\gamma^{(d)}\right)^2 + \mu_\gamma^{(d)}\left(\sigma_\gamma^{(t-1)}\right)^2}{\left(\sigma_\gamma^{(t-1)}\right)^2 + \left(\sigma_\gamma^{(d)}\right)^2} \\
  \sigma_\gamma^{(t)} & = \sqrt{\frac{\left(\sigma_\gamma^{(t-1)}\right)^2\left(\sigma_\gamma^{(d)}\right)^2}{\left(\sigma_\gamma^{(t-1)}\right)^2 + \left(\sigma_\gamma^{(d)}\right)^2}}
\end{align}
where the superscripts $(t)$ and $(t-1)$ refer to the inferred values for the current and previous trials, respectively, and $(d)$ to the parameters in \eref{eqn:data-dist}.

\subsection{Fixed-$\sigma$ model}
In this model, the force at each trial is assumed to be sampled from a normal
distribution and the mean of this distribution is inferred from the
observations, while the standard deviation is assumed to be known. The mean of
the force is assumed to also be normal with unknown mean and standard
deviation:
\begin{align}
  p(\gamma) &= N(\mu_t, \sigma) \\
  p(\mu_t) &= N(\mu_mu, \sigma_\mu)
\end{align}
In this manner, the uncertainty for this agent is at the level of the estimate
of the parameters of the force field, as opposed to being directly on the
force. Because the priors for $\mu_\mu$ are normal and the standard deviation
$\sigma$ (not $\sigma_mu$) is known, the update equations are given by:
\begin{align}
  \mu_\mu &= \frac{\sigma^2}{\sigma^2 + \sigma_\mu^2}\mu_\mu + \frac{\sigma_\mu^2}{\sigma^2 + \sigma_\mu^2} x \\
  \sigma_\mu &= \frac{\sigma^2\sigma_\mu^2}{\sigma^2 + \sigma_\mu^2}
\end{align}
where $x$ is the observation.

\subsection{Full uncertainty}
In this version, we will assume that for each context, the agent's belief over
the magnitude of the force is given by a normal distribution:
\begin{equation}
p(\gamma) = N(\mu_f, \sigma_f) \label{eqn:data-dist}
\end{equation}
where $\mu_f$ and $\sigma_f$ are parameters to be estimated at each
trial. Before the experiment begins, the agent will have a prior distribution
over these parameters. A standard Bayesian approach is to choose NormalGamma
priors:
\begin{equation}
p(\mu_f, \sigma_f) = NG(\mu_0, \nu_0, \alpha_0, \beta_0)
\end{equation}
where $\mu_0, \nu_0, \alpha_0$ and $\beta_0$ are free hyperparameters of the
model. Because participants are told that there are three contexts (baseline,
left and right), we assume the following hyperparameter values:
\begin{align} \mu_f^{(0)} &= (0, -1, 1) \\
  \nu_f^{(0)} &= 0.1 \\
  \alpha_f^{(0)} &= 1 \\
  \beta_f^{(0)} &= 0.5
\end{align}
Note that $\mu_f^{(0)}$ is different for each context and is thus
provided as a vector $(0, -1, 1)$ representing the baseline, left and right
contexts, respectively. For all other hyperparameters, all contexts have the
same value. These values ensure that the prior over the hyperparameters
$(\mu_f, \sigma_f)$ has the mode at $(\mu_f^{(0)}, 1)$. We will show that the
choice of these hyperparameters affects learning rate, but not the final
learned magnitude \todo{Check this}.

@@
